Supplemental Information B: Full Annotated Analysis and Reproducible Code

### 2022-01-14 Methods-and-Results.tex
meta General LM added
descr tables wide

We again started with a general test of the contact hypothesis, as in
Studies 1 and 2. We found no significant correlation between intergroup
contact and average outgroup attitudes --- for neither the total number
of outgroup interactions nor the number of measurement beeps with an
interaction (\textit{r} \textless{} 0.13, \textit{p} \textgreater{}
0.291). However, we did find a significant correlation between the
participants' Average Interaction Quality and their Average Outgroup
Attitudes (\textit{r} = 0.25, \textit{p} = 0.036). When considering the
number of interactions and average interaction quality jointly in a
linear regression, we only find a main effect of average perceived
interaction quality on predict outgroup attitudes (\textit{b} = 0.36,
\textit{t}(67) = 2.22, \textit{p} = 0.030, \(\eta_p^2\) = 0.07) but no
significant effect of the number of contacts nor an interaction effect.
This result mirrors that of study one but is inconsistent with the
second study. Additionally in a multilevel regression to check for
contemporaneous effects within the participants, we found that having an
outgroup interaction was indeed associated with significantly more
positive outgroup attitudes within the participants (random slopes
model; \textit{b} = 5.63, \textit{t}(3834) = 6.61, \textit{p}
\textless{} .001, \textit{95\%CI}{[}3.97, 7.29{]}), even after
controlling for having an interaction with a non-Dutch (which did not
relate to outgroup attitudes independently; For full results see Online
Supplementary Materials B). Thus, in our third data set we again found
mixed results, outgroup contacts are not significantly related to
outgroup attitudes on aggregate between participants and this effect is
also not suppressed by a average interaction quality (i.e., no
interaction effect). However, while the aggregate data is partly
inconsistent with the first two studies, the within person
contemporaneous effect of intergroup contact is consistent across all
three studies.

@@

To compare the models using either Allport's conditions or the core need
fulfillment to predict outgroup attitudes, we first assess relative
model performance indices (i.e., Akaike information criterion, and
Bayesian information criterion), and then consider the two predictors in
a joint model to see whether the two approaches predict the same
variance in outgroup attitudes. When comparing the model selection
indices we find that the fulfillment of the situation core need, indeed
performs slightly better than the model using Allport's conditions
(\(AIC_{CoreNeed}\) 12632.02 \textless{} 12651.59 \(AIC_{Allport}\), and
\(BIC_{CoreNeed}\) 12664.55 \textless{} 12684.12 \(BIC_{Allport}\)).
Additionally, when considering the predictors jointly, we find that both
significantly predict outgroup attitudes with similar sized regression
parameters (random slopes model; Allport's Conditions: \textit{b} =
0.16, \textit{t}(1600) = 4.92, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.09, 0.24{]}, Core Need: \textit{b} = 0.14,
\textit{t}(1600) = 3.85, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.08, 0.17{]}). This indicates that, although both
Allport's conditions and the core need fulfillment seem to (in part)
work through perceived interaction quality, they explain different
aspects of the variance in outgroup attitudes and do not constitute one
another.

@@

We again checked for alternative models to the core need fulfillment.
First, when considering generalized situational core need fulfillment
together with whether an intergroup contact took place, we find that
there is no significant main effect of core need fulfillment (random
slopes model; \textit{b} = 0.03, \textit{t}(3835) = 1.51, \textit{p} =
0.131, \textit{95\%CI}{[}-0.01, 0.06{]}) but a stronger interaction
effect of core need fulfillment and outgroup contact (\textit{b} = 0.17,
\textit{t}(3835) = 7.32, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.12, 0.21{]}). Together with a significant main
effect of having an outgroup contact (\textit{b} = 5.41,
\textit{t}(3835) = 6.36, \textit{p} \textless{} .001,
\textit{95\%CI}{[}3.74, 7.08{]}), this indicates that it is not key need
fulfillment in general --- but key need fulfillment during an outgroup
contact that predicts more positive outgroup attitudes. This finding is
consistent with the results of the previous studies.

In a final step we again checked whether during the interaction the core
situational need remains a meaningful predictor even when taking other
fundamental psychological needs into account. We find that the core need
adds significantly above a model with only the self determination theory
needs (random slopes models; \(\chi^2\)(6, \textit{N} = 70) = 100.20,
\textit{p} \textless{} .001). We find that the core need explains the
most variance in outgroup attitudes after an outgroup contact
(\textit{b} = 0.15, \textit{t}(1598) = 3.88, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.07, 0.22{]}). When compared to the model with
only the SDT needs, the core need fulfillment flexibly takes on some of
the explained variance of all of the three fundamental needs. However,
similar to the previous study, in this large sample relatedness
(\textit{b} = 0.05, \textit{t}(1598) = 3.31, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.02, 0.09{]}), competence (\textit{b} = 0.06,
\textit{t}(1598) = 2.73, \textit{p} = 0.006, \textit{95\%CI}{[}0.02,
0.10{]}) and autonomy (\textit{b} = 0.04, \textit{t}(1598) = 2.12,
\textit{p} = 0.034, \textit{95\%CI}{[}-0.01, 0.08{]}) each also
predicted positive outgroup attitudes independently. However, the
regression coefficient is three times as large for the core need
fulfillment (with all scaling being equal). For full results see Online
Supplementary Information B.

\input{Tables/descrFullLong}
\input{Tables/descrOutLong}

\input{Tables/mdlContactGeneralLong}

\input{Tables/mdlTheorylong}

\input{Tables/mdlRobustlong}

>>>>

As in Studies 1 and 2, we first conducted a general contact hypothesis
test. We found no correlation between intergroup contact and average
outgroup attitudes --- for neither the total number of outgroup
interactions nor the number of measurement beeps with an interaction
(\textit{r} \textless{} 0.13, \textit{p} \textgreater{} 0.291). However,
we did find a significant correlation between the participants' Average
Interaction Quality and their Average Outgroup Attitudes (\textit{r} =
0.25, \textit{p} = 0.036). When considering the number of interactions
and average interaction quality jointly in a linear regression, we only
found a main effect of average perceived interaction quality on outgroup
attitudes (\textit{b} = 0.36, \textit{t}(67) = 2.22, \textit{p} = 0.030,
\(\eta_p^2\) = 0.07) but no significant effect of the number of contacts
nor an interaction effect. This result mirrors that of study one but is
inconsistent with the second study. Additionally in a multilevel
regression to check for contemporaneous within person effects, we found
that having an outgroup interaction was indeed associated with
significantly more positive outgroup attitudes within the participants
(random slopes model; \textit{b} = 5.63, \textit{t}(3834) = 6.61,
\textit{p} \textless{} .001, \textit{95\%CI}{[}3.97, 7.29{]}), even
after controlling for having an interaction with a non-Dutch (which did
not relate to outgroup attitudes independently; for full results see
Online Supplementary Materials B). Thus, in our third data set we again
found mixed results. Outgroup contacts were on average not significantly
related to outgroup attitudes on a between participant level and this
effect was also not suppressed by an average interaction quality (i.e.,
no interaction effect). However, while the aggregate data was partly
inconsistent with the first two studies, the within person
contemporaneous effect of intergroup contact was consistent across all
three studies.

@@

To test wether Allport's conditions or the core need fulfillment are
better at predicting outgroup attitudes, we first assess relative model
performance indices (i.e., Akaike information criterion, and Bayesian
information criterion), and then consider the two predictors in a joint
model to see whether the two approaches predict the same variance in
outgroup attitudes. When comparing the model selection indices we find
that the fulfillment of the situational core need, indeed performs
slightly better than the model using Allport's conditions
(\(AIC_{CoreNeed}\) 12632.02 \textless{} 12651.59 \(AIC_{Allport}\), and
\(BIC_{CoreNeed}\) 12664.55 \textless{} 12684.12 \(BIC_{Allport}\)).
Additionally, when considering the predictors jointly, we find that both
significantly predict outgroup attitudes with similar sized regression
parameters (random slopes model; Allport's Conditions: \textit{b} =
0.16, \textit{t}(1600) = 4.92, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.09, 0.24{]}, Core Need: \textit{b} = 0.14,
\textit{t}(1600) = 3.85, \textit{p} \textless{} .001,
\textit{95\%CI}{[}0.08, 0.17{]}). This indicates that, although both
Allport's conditions and the core need fulfillment seem to (in part)
work through perceived interaction quality, they explain different
aspects of the variance in outgroup attitudes and do not constitute one
another.

@@

We again checked for alternative models. First, when considering
generalized situational core need fulfillment together with whether an
intergroup contact took place, we find that there is no significant main
effect of core need fulfillment (random slopes model; \textit{b} = 0.03,
\textit{t}(3835) = 1.51, \textit{p} = 0.131, \textit{95\%CI}{[}-0.01,
0.06{]}) but a stronger interaction effect of core need fulfillment and
outgroup contact (\textit{b} = 0.17, \textit{t}(3835) = 7.32, \textit{p}
\textless{} .001, \textit{95\%CI}{[}0.12, 0.21{]}). Together with a
significant main effect of having an outgroup contact (\textit{b} =
5.41, \textit{t}(3835) = 6.36, \textit{p} \textless{} .001,
\textit{95\%CI}{[}3.74, 7.08{]}), this indicates that it is not key need
fulfillment in general --- but key need fulfillment during an outgroup
contact that predicts more positive outgroup attitudes. This finding is
consistent with the results of the previous studies.

In a final step we again checked whether during the interaction the core
situational need remains a meaningful predictor even when taking other
fundamental psychological needs into account. We find that the core need
adds significantly above a model with only the self determination theory
needs (random slopes models; \(\chi^2\)(6, \textit{N} = 70) = 100.20,
\textit{p} \textless{} .001). We find that the core need explains the
most variance in outgroup attitudes after an outgroup contact
(\textit{b} = 0.15, \textit{t}(1598) = 3.88, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.07, 0.22{]}). When compared to the model with
only the SDT needs, the core need fulfillment flexibly takes on some of
the explained variance of all of the three fundamental needs. However,
similar to the previous study, in this large sample relatedness
(\textit{b} = 0.05, \textit{t}(1598) = 3.31, \textit{p} \textless{}
.001, \textit{95\%CI}{[}0.02, 0.09{]}), competence (\textit{b} = 0.06,
\textit{t}(1598) = 2.73, \textit{p} = 0.006, \textit{95\%CI}{[}0.02,
0.10{]}) and autonomy (\textit{b} = 0.04, \textit{t}(1598) = 2.12,
\textit{p} = 0.034, \textit{95\%CI}{[}-0.01, 0.08{]}) each also
predicted positive outgroup attitudes independently. This being said,
the regression coefficient for the core need was three times larger
(with all scaling being equal). For full results see Online
Supplementary Information B.

\input{Tables/descrFullWide}
\input{Tables/descrOutWide}

\input{Tables/mdlContactGeneralLong}
\input{Tables/mdlTheorylong}
\input{Tables/mdlRobustlong}
